---
title: "Viz part 1"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
```

---set the size and scale of plots in md shown up 

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

Get the data for plotting today.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Let's make a plot!

```{r}
ggplot(weather_df,aes(x = tmin, y = tmax)) +
 geom_point()
```

Pipes and stuff

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()


ggp_nyc_weather =
  weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```


## Fancy plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + # if color here, then work for the whole command
  geom_point(aes(color = name, alpha = 0.3))+ # if put color here then no color for smooth,alpha uses to change opqueness and solidness
  geom_smooth(se = FALSE) #by add standard error=FALSE, get rid of error bars,grey area
```


Plot with facets

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE)+
  facet_grid(. ~ name) #one pannel in each name variable, central_park...; .put everything in one row, and then separate variable according to name variable according to columns;if name ~.,then 3 rows 1 column
```

Let's try a different plot.temps are boring

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name))+
  geom_point(aes(size = prcp),alpha = 0.3) + #higher prcp variable have bigger points
  geom_smooth()+
  facet_grid(. ~ name)
```


try assigning specific color(#not usually do so, usually assign by default as the previous chunk)

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x = date, y = tmax))+ #doing color=blue does not work here
  geom_point(color="blue",alpha=0.7, size=0.5)  #assign color in geom_point function
```


```{r}
weather_df |> 
  filter(name != "CentralPark_NY") |> 
  ggplot(aes(x = date, y = tmax,color = name))+ 
  geom_point(alpha=0.7, size=0.5)  #size&alpha we usually set?
```

hex plot
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y =tmax))+
  geom_hex() #install a package doing density plots;if you have many data over 1000,try do this density plot,each plot occupies some space
```

Line plot
```{r}
weather_df |> 
  filter(name == "Molokai_HI") |> 
  ggplot(aes(x=date, y =tmax)) +
  geom_line(alpha=0.5) +
  geom_point(size=0.5)
```


## univariate plotting
only look at one variable
easy way to do so histogram

```{r}
ggplot(weather_df, aes(x = tmax,fill= name))+ #if use color,outside the bar;here we use fill, fill in the bar with color
  geom_histogram(position = "dodge") #it means bars should avoid each other, not on top of each other
```

let's use a density plot (smooth histogram)

```{r}
ggplot(weather_df, aes(x = tmax,fill= name)) +
    geom_density(alpha = 0.3, adjust = 0.75) #adjust-- make smoother
```

using boxplots!

```{r}
ggplot(weather_df, aes(y = tmax, x=name))+
  geom_boxplot()
```

violin plots?(all info same as density plot, but order vertically)

```{r}
ggplot(weather_df, aes(y = tmax, x=name))+
  geom_violin()
```

ridge plot (crate 3 density, but seprate not overlapping on each top)

```{r}
ggplot(weather_df, aes(x = tmax, y=name))+
  geom_density_ridges()
```


## saving and embedding plots

```{r}
ggp_weather =
  weather_df |> 
  ggplot(aes(x=tmin,y=tmax))+
  geom_point()

ggp_weather

ggsave("resulted plots/ggp_weather.pdf", ggp_weather)
```

change the scale of plot (#for md shown after git)
```{r, fig.width=12, fig.height==12}
ggp_weather
```

(#fig.asp=0.6, a little bit wider than its height)
```{r, fig.width=12, fig.asp==6} 
ggp_weather
```
